A mechanization of semi-simplicial and semi-cubical sets in the calculus of inductive constructions

### 2022-10-18 paper/paper.tex
paper: provide nice pic for intuition; X_0, X_1, ...

% Some abbreviations
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Type}{\textsf{Type}}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

@@

\subsection{Intuition for our formal construction}
To assign $X_0, X_1, X_2, \ldots$ in the above indexed representation types, we need to exhibit a recursive definition over our building blocks, which we will call $\framep$, $\layer$, and $\filler$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), a $\filler$ is a filled $\framep$; both of them can be decomposed into $\layer$.

As a first step, we assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. The question is then to find the recursive recipe defining $\fullframe[n]$. First, $\fullframe[n]$ depends on all $X_i$ upto $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. We shall give the full definition of $\fullframe[n]$ shortly, but first, assuming $\fullframe[n]$ is defined, we are able to take the coinductive limit of the $\X[][<n][]$, as shown in \ref{tab:coind}, therefore fully characterizing the indexed definition of $\nu$-sets.

>>>>

% Some abbreviations
\newcommand{\tl}{\ensuremath{\mathsf{tl}}}
\newcommand{\hd}{\ensuremath{\mathsf{hd}}}
\newcommand{\imp}{\rightarrow}
\newcommand{\overright}[1]{\overrightarrow{#1}}
\renewcommand{\D}{D}
\newcommand{\hdD}{D.1}
\newcommand{\tlD}{D.2}
\renewcommand{\d}{d}
\renewcommand{\E}{E}
\newcommand{\ap}{\mathsf{ap}}
\renewcommand{\l}{l}
\renewcommand{\c}{c}
\newcommand{\pair}[2]{#1, #2}
\newcommand{\Type}{\textsf{Type}}
\newcommand{\Dom}{\textsf{Dom}}
\newcommand{\UIP}{\textsf{UIP}}

@@

\subsection{Intuition for our formal construction}
\begin{notation}[\Sigma]
  In the prose that follows, we assume the existence of $\Sigma$-types with projections $\hd$ and $\tl$.
\end{notation}

To assign $X_0, X_1, X_2, \ldots$ in the above indexed representation types, we need to exhibit a recursive definition over our building blocks, which we will call $\framep$, $\layer$, and $\filler$. A $\framep$ is a boundary of a standard form (simplex, cube, etc.), a $\filler$ is a filled $\framep$; both of them can be decomposed into $\layer$.

As a first step, we assign every $X_n$ the type $\fullframe[n] \rightarrow \U$ uniformly, applying the isomorphism between $A \rightarrow B \rightarrow C$ and $A \times B \rightarrow C$, or between $\Pi a : A, (B a \rightarrow C)$ and $(\Sigma a : A, B a) \rightarrow C$. The question is then to find the recursive recipe defining $\fullframe[n]$. First, $\fullframe[n]$ depends on all $X_i$ upto $n - 1$. So, we need to package up $X_i$, for $i < n$, into a type, which we call $\X[][<n][]$. This allows us to give the type $\X[][<n][] \rightarrow \U$ to $\fullframe[n]$. Then, for $D: \X[][<n][]$, representing an initial prefix of $X_0, X_1, \ldots X_{n - 1}$, the indexed set $X_n$ has type $\fullframe[n](D) \rightarrow \U$. We shall give the full definition of $\fullframe[n]$ shortly, but first, assuming $\fullframe[n]$ is defined, we are able to take the coinductive limit of the $\X[][<n][]$, as shown in \ref{tab:coind}, therefore fully characterizing the indexed definition of $\nu$-sets. Let us illustrate how we will build $\fullframe[n]$ by refining types of $X_0, X_1, \ldots, X_{n - 1}$ so that they have the form $\fullframe[n] \rightarrow \U$. To begin, let us set $\fullframe[0] = \unittype$, so that the type $\U$ of $X_0$ can be equivalently formulated as $\unittype \rightarrow \U$. The rest of the construction follows.

\begin{equation*}
  \begin{array}{llll}
    X_0                & :              & \unittype                                   & \rightarrow \U \\
    X_1                & :              & \Sigma \kstar: \unittype. \left(
    \begin{array}{c}
        X_0 \kstar
        \\ \times \\
        X_0 \kstar
      \end{array}\right) & \rightarrow \U                                                                \\
    X_2                & :              & \Sigma a: (\Sigma \kstar: \unittype. \left(
    \begin{array}{c}
        \Sigma c: \left(
        \begin{array}{c}X_0 \kstar
            \\ \times \\
            X_0 \kstar
          \end{array}
        \right). (X_1 (\kstar, c)
        \\ \times \\
        \Sigma c: \left(
        \begin{array}{l}
            X_0 \kstar
            \\ \times \\
            X_0 \kstar
          \end{array}
        \right). (X_1 (\kstar, c)
      \end{array}
    \right).\left(
    \begin{array}{c}
        X_1 \left(\kstar, \left(
          \begin{array}{c}
              a.\tl.0.\hd.0, \\
              a.\tl.1.\hd.0
            \end{array}
          \right)\right)
        \\ \times \\
        X_1 \left(\kstar, \left(
          \begin{array}{c}
              a.\tl.0.\hd.1, \\
              a.\tl.1.\hd.1
            \end{array}
          \right)\right)
      \end{array}
    \right)
                       & \rightarrow \U                                                                \\
    \ldots
  \end{array}
\end{equation*}
